Multi-task learning in Computer Vision

### 2020-04-01 1-Introduction.tex
Seminar done

>>>>

Here we will take a look at transfer learning and how it relates to multi-task learning.
Then we will define many aspects of multi-task learning and gain an understanding of the ways that representations can be shared in neural networks.
Finally, we will look at some examples that demonstrate both the benefits and the costs of solving a multi-task learning problem.

### 2020-04-01 3-Image-classification.tex
Seminar done

To give a formal definition of transfer learning, we will follow the definitions provided in \citep{transferSurvey2010}. Let $\mathcal{D}$ be a domain, which consists of a feature space $\mathcal{X}$ and a marginal probability distribution P($\mathcal{X}$). For a given domain, a task $\mathcal{T}$ = {$\mathcal{Y}$, f$\mathcal(X)$} consist of a label space $\mathcal{Y}$ for the inputs and of a predictive function $f(x)$ which produces predictions for all pairs ${x_i, y_i}$ where $x_i \in \mathcal{X}$ and $y_i \in \mathcal{Y}$. Given a source domain $\mathcal{D}_S$, a source task $\mathcal{T}_S$, a target domain $\mathcal{D}_T$ and a target task $\mathcal{T}_T$, where target and source are disjoint, transfer learning tries to improve the performance of $f_T(x)$ using $\mathcal{D}_S$ and $\mathcal{T}_S$.

>>>>

To give a formal definition of transfer learning, we will follow the definitions provided in \citep{transferSurvey2010}. Let $\mathcal{D}$ be a domain, which consists of a feature space $\mathcal{X}$ and a marginal probability distribution P($\mathcal{X}$). For a given domain, a task $\mathcal{T}$ = \{$\mathcal{Y}, f\mathcal(X)$\} consist of a label space $\mathcal{Y}$ for the inputs and of a predictive function $f(x)$ which produces predictions for all pairs \{$x_i, y_i$\} where $x_i \in \mathcal{X}$ and $y_i \in \mathcal{Y}$. Given a source domain $\mathcal{D}_S$, a source task $\mathcal{T}_S$, a target domain $\mathcal{D}_T$ and a target task $\mathcal{T}_T$, where target and source are disjoint, transfer learning tries to improve the performance of $f_T(x)$ using $\mathcal{D}_S$ and $\mathcal{T}_S$.

### 2020-04-01 5-Multitask.tex
Seminar done

Once the loss function is specified, the actual training is quite similar to training a Single-task model.
The one new thing to consider in a Multi-task setting is the sampling ratio of the different tasks as some tasks can be easier to learn or contain significantly more or less data compared to other tasks.
Like in regular training, the models are trained one batch at a time.
In a Multi-task setting, we have datasets for each of the tasks from which we can pick at a time.
Now, we can define an epoch in the training process as several batches with an expected value of $\sum_i{ \dfrac{|X_i| \alpha_i}{B_i}}$ batches over $i$ tasks, where $|X_i|$ is the training size, $B_i$ is the batch size, and $\alpha_i$ is the scaling factor for each task.
Iterating through all batches can be done in a round-robin fashion,  going through all batches of each data set at a time or randomly sampling the sets with probabilities respective to their magnitude and scaling factor.

\section{Latent image representations}
For a Multi-task model, the optimal situation is such, where all classifiers would use a single shared latent representation of the image to predict all attributes.
This representation is an Image Embedding, that would capture all valuable information of the image for our tasks.
The embedding could be the vector represented by the final layer of an ImageNet classifier flattened to some n-dimensional vector that then solves all the tasks instead of just a single task.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{imgs/sharedBackbone.png}
    \caption{Example of partly and completely shared backbones. Figure from \citep{visualPerson}.\label{fig:params}}
\end{figure}
However, such universal image representation can be challenging or impossible to learn, and instead, only a part of the network is shared between the tasks.
Here, this shared structure is called the backbone of the Multi-task classifier.
Often the backbone is in the form of some ImageNet classifier architecture, and it can be the entire classifier or some layers up to some arbitrary limit.
As the amount of layers shared is picked for each task separately, some tasks can share the entire classifier as a backbone, and others might only share some amount of layer that produces desirable results.

Besides just deciding the network branching, each task requires an independent head that outputs prediction for that task.
These task heads are similar to those used in transfer learning, where only a single task is solved using the ImageNet backbone.
Especially in a Multi-task setting, these heads can be more than just a single linear layer calculating the softmax over all classes.
A single head can contain several linear layers, dropout layers, batch normalization layers, or it could be a Recurrent Neural Network creating a caption for the image embedding, depending on the task at hand.
In Figure 3.1, we can see an example of the flexibility of the sharing as the model can use a completely shared backbone or a branching backbone, which has a unique head for each task, solving the person re-identification and classification tasks.

\section{Benefits}
Depending on how Multi-task learning is applied, it can provide a multitude of benefits to the model.
Many of these benefits stem from the fact that the model gets to see more data, and the various tasks can make it easier to find useful features.
Different kinds of data sets have different kinds of noise, learning multiple tasks makes it easier to distinguish which features are beneficial and detrimental some features may be difficult to learn on a dataset but can be useful and borrowed from another and learning multiple tasks forces the model to not overfit on one of them \citep{ruderOverview}.

@@

\section{Hard parameter sharing}
A hard parameter sharing setting is one where the same convolutional weights in the intermediate layers solve multiple tasks, like in Figure 3.1.
Hard parameter sharing is the primary way of reducing total model size and inference time when solving multiple tasks simultaneously.
While sharing the weights can provide many benefits to the end model, as seen in the previous chapter, it comes at the cost of an increased number of tunable parameters when training the models.
As we have previously seen, Multi-task models introduce many new constant factors to the training process beyond the normal hyperparameters for learning rate, dropout rate, and others.
These include the hyperparameters for weighing the loss functions and determining the task sampling ratios, but also the expensive to evaluate which tasks should share representations and how much should be shared.

@@

This kind of sharing does not scale very well as the number of tasks increases since calculating the relation between various tasks is quadratic.
Since the joining logic requires extra parameters, it means there is even more to learn than just learning all the tasks separately.
As soft sharing does not provide many of the benefits that are gained by hard sharing the parameters, it is not as popular, but still used in various cases.
Learning the soft connections between layers does not reduce the parameters in the model.
However, it is useful in adding some extra information between the tasks to gain extra performance by utilizing the latent representations between them in various ways.

\section{Special multi-task learning techniques}
Show how an attention layer can be added to augment the performance in various tasks in a multi-task setting \citep{multiTaskAttention}, especially when the tasks are correlated \citep{multiTaskWeather} \citep{weatherNet}. Maybe some others as well.

@@

It would make sense that the results of the Taskonomy would be easily transferable to the Multi-task setting.
When evaluating the Multi-task vs. Transfer affinity, it turns out that they are negatively correlated, at least in the case of the five tasks that were the focus of the experiments in \citep{whichTasks}.
Based on this observation, it can be beneficial to train non-related tasks together.
The authors suggest that the different tasks act as a good form of regularization.
It could also be that some of the learned features work very well for the other task, but can't be easily learned with the dataset of the other task and vice versa.

>>>>

Once the loss function is specified, the actual training is quite similar to training a Single-task model.
The one new thing to consider in a Multi-task setting is the sampling ratio of the different tasks as some tasks can be easier to learn or contain significantly more or less data compared to other tasks.
Like in regular training, the models are trained one batch at a time.
In a Multi-task setting, we have datasets for each of the tasks from which we can pick at a time.
Now, we can define an epoch in the training process as the total number of batches from all tasks.
We can define this as $\sum_i{ \dfrac{|X_i| \alpha_i}{B_i}}$ batches over $i$ tasks, where $|X_i|$ is the training set size, $B_i$ is the batch size, and $\alpha_i$ is the scaling factor for each task, telling how important the specific task is.
Iterating through all batches can be done in a round-robin fashion, by going through all batches of each data set at a time or randomly sampling the sets with probabilities respective to their magnitude and scaling factor.

\section{Latent image representations}
For a Multi-task model, the optimal situation is such, where all classifiers would use a single shared latent representation of the image to predict all attributes.
This representation is an Image Embedding, that would capture all valuable information of the image for our tasks.
The embedding could be the vector represented by the final layer of an ImageNet classifier flattened to some n-dimensional vector that then solves all the tasks instead of just a single task.
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{imgs/sharedBackbone.png}
    \caption{Example of partly and completely shared backbones. Figure from \citep{visualPerson}.\label{fig:params}}
\end{figure}
However, such a universal image representation can be challenging or impossible to learn, and instead, only a part of the network is shared between the tasks.
Here, this shared structure is called the backbone of the Multi-task classifier.
Often the backbone is in the form of some ImageNet classifier architecture, and it can be the entire classifier or some layers up to some arbitrary limit.
As the amount of layers shared is picked for each task separately, some tasks can share the entire classifier as a backbone, and others might only share some amount of layer that produces desirable results.

Besides just deciding the network branching, each task requires an independent head that outputs prediction for that task.
These task heads are similar to those used in transfer learning, where only a single task is solved using the ImageNet backbone.
Especially in a Multi-task setting, these heads can be more than just a single linear layer calculating the softmax over all classes.
A single head can contain any combination of several linear layers, dropout layers, batch normalization layers, convolutional layers, or it could be a Recurrent Neural Network creating a caption for the image embedding, depending on the task at hand.
In Figure 3.1, we can see an example of the flexibility of the sharing as the model can use a completely shared backbone or a branching backbone, which has a unique head for each task, solving the person re-identification and classification tasks.

\section{Benefits}
Depending on how and where Multi-task learning is applied, it can provide a multitude of benefits to the model.
Many of these benefits stem from the fact that the model gets to see more data, and the various tasks can make it easier to find useful features.
Different kinds of data sets have different kinds of noise, learning multiple tasks makes it easier to distinguish which features are beneficial and detrimental some features may be difficult to learn on a dataset but can be useful and borrowed from another and learning multiple tasks forces the model to not overfit on one of them \citep{ruderOverview}.

@@

\section{Hard parameter sharing}
A hard parameter sharing setting is one where the same convolutional weights in the intermediate layers solve multiple tasks, like in Figure 3.1.
Hard parameter sharing is the primary way of reducing total model size and inference time when solving multiple tasks simultaneously.
While sharing the weights can provide many benefits to the end model, as seen in the previous chapter, it comes at the cost of an increased number of tunable parameters when training the models.
As we have previously seen, Multi-task models introduce new constant factors to the training process beyond the normal hyperparameters for learning rate, dropout rate, and others.
These include the hyperparameters for weighing the loss functions and determining the task sampling ratios, but also the expensive to evaluate architectural decision of which tasks should share representations and how much should be shared.

@@

This kind of sharing does not scale very well as the number of tasks increases since calculating the relation between various tasks is quadratic.
Since the joining logic requires extra parameters, it means there is even more to learn than just learning all the tasks separately.
As soft sharing does not provide many of the benefits that are gained by hard sharing the parameters, it is not as popular, but still used in various cases.
However, it is useful in adding some extra information between the tasks to gain extra performance by utilizing the latent representations between them in various ways.
Also, in the basic case of using a soft parameter share to constraint, the layers can improve the generalizability of the model.

\section{Attention augmented Multi-task learning}
In a Multi-task setting, the model ends up having a representation of the image from each task's point of view, and it is possible to use attention \citep{attention} to augment the predictions.
Especially in the case where the tasks are heavily correlated, this can be quite useful.
For example, when doing weather recognition, it would make sense that some of the features are correlated, and this correlation can be used by adding an attention layer between the different tasks, like in \citep{cnn-rnn}.
The Multi-Task Attention Network (MTAN) \citep{multiTaskAttention} is an example of a more advanced application of attention in Multi-task models.
In the MTAN, each task gets its own attention module, and they are used to determine correlations of the tasks at specific layers of the network using the attention operations.
It is similar to cross-stitching networks in that each task has to learn its attention module features.
However, there is only a single backbone for the shared features, and the attention modules are responsible for producing the task level outputs.
The number of parameters is still significantly reduced as the task specific-parts are not complete ImageNet classifiers, resulting in a significantly more efficient and accurate classifier for image segmentation.

@@

It would make sense that the results of the Taskonomy would be easily transferable to the Multi-task setting.
When evaluating the Multi-task vs. Transfer affinity, it turns out that they are negatively correlated, at least in the case of the five tasks that were the focus of the experiments in \citep{whichTasks}.
Based on this observation, it can be beneficial to train some non-related tasks together.
The authors suggest that the different tasks act as a good form of regularization as the models need to generalize to multiple types of inputs.
It could also be that some of the learned features work very well for the other task, but can't be easily learned with the dataset of the other task and vice versa.
These results are aligned with the empirical results that we saw when we looked at the compatibility of classifiers using hard weight sharing.
Currently, it seems not to be possible to determine the compatibility of different tasks purely on a theoretical grounding.
Instead, experiments need to be run, and the results of pairs of tasks do not always generalize to a set of all other tasks when using shared representations in models \citep{visualPerson}.

### 2020-04-01 HY-CS-main.tex
Seminar done

%% For seminar papers and such, cover page and abstract
%% requires these three basic items of information.
%% Label as needed and remove comment marks.
%%\level{Seminaariraportti}
%%\programme{Tietojenkäsittelytieteen andidaattiohjelma}
%%\subject{Seminaarisarjan nimi}

@@

% BSc instructions
%\include{bsc_finnish_contents}
%\include{bsc_english_contents}
% MSc instructions
%\include{msc_finnish_contents}
\include{1-Introduction}
\include{3-Image-classification}
\include{5-Multitask}
\include{7-Experiments}
\include{8-Conclusions}
% \include{msc_english_contents}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\backmatter
\begin{appendices}

%\input{instructions_english}

\appendix{Sample Appendix\label{appendix:model}}
usually starts on its own page, with the name and number of the appendix at the top. 
The appendices here are just models of the table of contents and the presentation. Each appendix
Each appendix is paginated separately.

In addition to complementing the main document, each appendix is also its own, independent entity.
This means that an appendix cannot be just an image or a piece of programming, but the appendix must explain its contents and meaning.

\end{appendices}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%% For seminar papers and such, cover page and abstract
%% requires these three basic items of information.
%% Label as needed and remove comment marks.
% \level{seminar}
% \programme{Tietojenkäsittelytieteen andidaattiohjelma}
% \subject{Seminaarisarjan nimi}

@@

% BSc instructions
%\include{bsc_finnish_contents}
%\include{bsc_english_contents}
% MSc instructions
%\include{msc_finnish_contents}
\include{1-Introduction}
\include{3-Image-classification}
\include{5-Multitask}
\include{8-Conclusions}
% \include{msc_english_contents}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\backmatter
% \begin{appendices}

%     %\input{instructions_english}

%     \appendix{Sample Appendix\label{appendix:model}}
%     usually starts on its own page, with the name and number of the appendix at the top.
%     The appendices here are just models of the table of contents and the presentation. Each appendix
%     Each appendix is paginated separately.

%     In addition to complementing the main document, each appendix is also its own, independent entity.
%     This means that an appendix cannot be just an image or a piece of programming, but the appendix must explain its contents and meaning.

% \end{appendices}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
